---
output: github_document
editor_options: 
  chunk_output_type: console
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# senadoRES

<!-- badges: start -->
[![R build status](https://github.com/llrs/senadoRES/workflows/R-CMD-check/badge.svg)](https://github.com/llrs/senadoRES/actions)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![Codecov test coverage](https://codecov.io/gh/llrs/senadoRES/branch/master/graph/badge.svg)](https://codecov.io/gh/llrs/senadoRES?branch=master)
<!-- badges: end -->

The goal of senadoRES is to provide information about the [Senate of Spain](https://en.wikipedia.org/wiki/Senate_of_Spain).
It uses the [Open Data](https://www.senado.es/web/relacionesciudadanos/datosabiertos/catalogodatos/index.html) available and some other XML files I found while exploring the website.

## Installation

You can install the released version of senadoRES with:


``` r
remotes::install_github("llrs/senadoRES")
```

## Example

To make it easier to work and show some data I'll use some packages:

```{r example}
library("senadoRES")
library("dplyr")
library("ggplot2")
```

### Senadores

This function return the members of the Senate:

```{r senadores}
ts <- senadores()
head(ts)
```


We can see the increase of females on the Senate along the legislatures:

```{r senadoras}
ts %>%
  group_by(legislatura) %>%
  count(sex) %>%
  mutate(total = sum(n)) %>%
  filter(!is.na(sex)) %>%
  mutate(ratio = n/total) %>%
  filter(sex != "male") %>% 
  ggplot() +
  geom_point(aes(legislatura, ratio), size = 5) +
  geom_hline(yintercept = 0.5, linetype = 2, col = "red") +
  scale_x_continuous(breaks = seq_len(15)) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1), 
                     breaks = seq(from = 0, to = 1, by = .1),
                     expand = expansion(add = c(0.01, 0.01)), limits = c(0, NA)) +
  theme_minimal() +
  labs(title = "Ratio of women", x  = "Legislatura", y = "% of women") +
  guides(col = FALSE, shape = FALSE) +
  theme(panel.grid.minor.x = element_blank())
```

Or see what has been done:

```{r boletin}
b <- boletin(boletin_csv(14, 1))
b[, c("FECHA", "DISP", "apartado", "subapartado", "TITULO", "document")]
```

So, on the first session of the 14th legislature they dealt with human resources.
We can check one of those documents, to retrieve the text with `documento`.
