---
output: github_document
editor_options: 
  chunk_output_type: console
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# senadoRES

<!-- badges: start -->
<!-- badges: end -->

The goal of senadoRES is to provide information about the [Senate of Spain](https://en.wikipedia.org/wiki/Senate_of_Spain).
It uses the [Open Data](https://www.senado.es/web/relacionesciudadanos/datosabiertos/catalogodatos/index.html) available and some other XML files I found while exploring the website.

## Installation

You can install the released version of senadoRES from [CRAN](https://CRAN.R-project.org) with:

``` r
install.packages("senadoRES")
```

or 

``` r
remotes::install_github("llrs/senadoRES")
```

## Example

To make it easier to work and show some data I'll use some packages:

```{r example}
library("senadoRES")
library("dplyr")
library("ggplot2")
```

### Senadores

This function return the members of the Senate:

```{r senadores}
ts <- senadores()
head(ts)
```


We can see the increase of females on the Senate along the legislatures:

```{r senadoras}
ts %>%
    group_by(legislatura) %>%
    count(sex) %>%
    mutate(total = sum(n)) %>%
    filter(!is.na(sex)) %>%
    mutate(ratio = n/total) %>%
    filter(sex != "male") %>% 
    ggplot() +
    geom_point(aes(legislatura, ratio, col = sex, shape = sex), size = 5) +
    geom_hline(yintercept = 0.5, linetype = 2, col = "red") +
    scale_x_continuous(breaks = seq_len(15)) +
    scale_y_continuous(labels = scales::percent_format(accuracy = 1), 
                       breaks = seq(from = 0, to = 0.7, by = .1),
                       expand = expansion(add = c(0, 0.01)), limits = c(0, NA)) +
    theme_minimal() +
    labs(title = "Ratio of women", x  = "Legislatura", y = "% of women") +
    guides(col = FALSE, shape = FALSE)
```


Or the change of the parties along the years:

```{r grupos}
ts %>% 
  group_by(legislatura) %>%
  count(grupoNombre) %>% 
  mutate(grupoNombre = case_when(legislatura == "4" ~ gsub("^G", "GP", grupoNombre),
                                 TRUE ~ grupoNombre)) %>% # Fix an obvious error
  ggplot() +
  geom_tile(aes(legislatura, 
                forcats::fct_reorder2(grupoNombre, legislatura != 0, -legislatura, .fun = sum)),
            fill = "black") +
  scale_x_continuous(breaks = seq_len(15)) +
  theme_minimal() +
  labs(title = "Grupos políticos", x  = "Legislatura", y = element_blank())
```

